---
title: "elo_fit"
author: "Christian Tillich"
date: "October 22, 2018"
output: md_document
---

```{r setup, include=FALSE}
library(reshape2)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
```

# Elo Convergence

```{r elo_dist, out.height = 250, out.width = 250}
df12 <- read.csv('../data/with_elo_k12.csv')
df20 <- read.csv('../data/with_elo_k20.csv')
df32 <- read.csv('../data/with_elo_k32.csv')

qplot(df12$elo_P1); qplot(df12$elo_P2)
qplot(df20$elo_P1); qplot(df20$elo_P2)
qplot(df32$elo_P1); qplot(df32$elo_P2)
```


# Objective

To assess the fit of our three elo models qualitatively. By assessing the individual narratives of some of the more high-profile players in SC2 in this time period, we can get the following:

1) Whether those stories match known observations at the time. 
2) Whether a larger/smaller k makes for a better model. 
3) What's an acceptable burn-in rate to exclude games where elo may not be representative of the player's skill. 



```{r elo_story}
elo_story <- function(player, df = df12){
  df %>% 
    filter(P1 == player | P2 == player) %>% 
    select(id, P1,P2,elo_P1,elo_P2) %>% 
    mutate(player_score = ifelse(P1 == player, elo_P1, elo_P2)) %>%
    select(id, winner = P1, loser = P2, elo = player_score)
}

```


## IdrA's Story

```{r idra_story}
elo_story('600_IdrA') %>% qplot(x = id, y = elo, data = .)
elo_story('600_IdrA', df20) %>% qplot(x = id, y = elo, data = .)
elo_story('600_IdrA', df32) %>% qplot(x = id, y = elo, data = .)
```

Note, Idra's story really captures the "long slump well I think" [](https://liquipedia.net/starcraft2/IdrA)


## HerO's Story

```{r hero_story}
elo_story('222_HerO') %>% qplot(x = id, y = elo, data = .)
elo_story('222_HerO', df20) %>% qplot(x = id, y = elo, data = .)
elo_story('222_HerO', df32) %>% qplot(x = id, y = elo, data = .)
```
HerO starts in 2011. 


## Mvp's Story

```{r mvp_story}
elo_story('590_Mvp') %>% qplot(x = id, y = elo, data = .)
elo_story('590_Mvp', df20) %>% qplot(x = id, y = elo, data = .)
elo_story('590_Mvp', df32) %>% qplot(x = id, y = elo, data = .)
```

## Life's

```{r mvp_story}
elo_story('2747_Life') %>% qplot(x = id, y = elo, data = .)
elo_story('2747_Life', df20) %>% qplot(x = id, y = elo, data = .)
elo_story('2747_Life', df32) %>% qplot(x = id, y = elo, data = .)
```

Something is incorrect here. Life has an elo well above 1500 but he doesn't come out when running code to identify the best players. Look into this. 

```{r}
df12 %>% 
  group_by(P1) %>% 
  summarise(high_score = mean(elo_P1)) %>% arrange(desc(high_score)) %>% 
  head
```