---
title: "elo_fit"
author: "Christian Tillich"
date: "October 22, 2018"
output: md_document
---

```{r setup, include=FALSE}
library(reshape2)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
```

# Elo Convergence

```{r elo_dist, out.height = 250, out.width = 250}
df12 <- read.csv('../data/with_elo_k12.csv')
df20 <- read.csv('../data/with_elo_k20.csv')
df32 <- read.csv('../data/with_elo_k32.csv')

qplot(df12$elo_P1); qplot(df12$elo_P2)
qplot(df20$elo_P1); qplot(df20$elo_P2)
qplot(df32$elo_P1); qplot(df32$elo_P2)
```

# Date QA

```{r}
df12 %>% 
  qplot(x = Date, y = id, geom = "boxplot", data = df12)
```


# Objective

To assess the fit of our three elo models qualitatively. By assessing the individual narratives of some of the more high-profile players in SC2 in this time period, we can get the following:

1) Whether those stories match known observations at the time. 
2) Whether a larger/smaller k makes for a better model. 
3) What's an acceptable burn-in rate to exclude games where elo may not be representative of the player's skill. 



```{r elo_story}
elo_story <- function(player, df = df12){
  df %>% 
    filter(P1 == player | P2 == player) %>% 
    select(id, P1,P2,elo_P1,elo_P2) %>% 
    mutate(player_score = ifelse(P1 == player, elo_P1, elo_P2)) %>%
    select(id, winner = P1, loser = P2, elo = player_score)
}

add_time_mark <- function(plot, pos, label, y = 950){
  plot +
    geom_vline(xintercept = pos) + 
    geom_text(x = pos + 800, y = y, label = label, angle = 90) + 
    coord_cartesian(ylim = c(y - 50, 1500))
}

tournament_search <- function(search_term, data = df12){
  data %>%
    filter(grepl(search_term, tolower(League))) %>%
    group_by(League) %>%
    summarise(first_game = min(id), last_game = max(id), start_date = min(Date), end_date = max(Date)) %>% 
    arrange(first_game)
}

best_elo <- function(df = df12) {
  getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
  }
  
  df %>% 
    {rbind(
        select(., player = P1, elo = elo_P1, race = Race_P1)
      , select(., player = P2, elo = elo_P2, race = Race_P2)
    )} %>% 
    group_by(player) %>% 
    summarise(max = max(elo), min = min(elo), most_played = getmode(race), count = n()) %>% 
    arrange(desc(max)) %>% as.data.frame
}
```


## IdrA's Story

```{r idra_story}
elo_story('600_IdrA') %>% qplot(x = id, y = elo, data = .) %>% 
  add_time_mark(849, "MLG Raliegh 2010") %>%
  add_time_mark(8598, "GSL Code S 2010") %>%
  add_time_mark(17425, "MLG Dallas 2011") %>%
  add_time_mark(36470, "MLG Raleigh 2011") %>%
  add_time_mark(40192, "IEM Guangzhou 2011") %>%
  add_time_mark(49987, "GSL Code S 2012") %>%
  add_time_mark(71768, "ASUS GSP Pro 2012")
elo_story('600_IdrA', df20) %>% qplot(x = id, y = elo, data = .)
elo_story('600_IdrA', df32) %>% qplot(x = id, y = elo, data = .)
```

IdrA is perhaps one of the most wildly popular and controversial SC2 players, 
and I included him in the QA sample because the narratives around his performance
suggest swings in behavior that our model should pick up on. And indeed it does.

Our data set starts the day after the SC2 release, with the Go4SC2 Cup, and quickly
moves into the 2010 MLG tournament. Idra did very well initially, taking 
first place in MLG 2010 and securing an offer with the Evil Geniuses, a North
American Starcraft 2 team. 

By summer 2011, however, IdrA starts showing more moderated performance, with 
some disappointing tournament results combined with a string of games that IdrA
conceded, despite having advantageous positions. There's a brief uptick in 
Fall 2011, but by Jan 2012 IdrA begins a period known as "the long slump".

The elo model captures this long slump period well, showing IdrA dropping almost
200 points betweeen Winter and Summer 2012. IdrA does manage to rebound ~125 
points after his low mark, but unfortunately our data set ends there, so we can't
comment further on his recovery. 

Our goal here is to ensure that the elo model follows the narrative arc of the 
player sufficiently, and IdrA has a dynamic story to tell over those two years
which the model captures quite nicely. However, we must take care to not place
too much interpretive weight on the model scores themselves at this point. Because
if SC2 was imbalanced against Zerg, we might expect that bias to get baked into elo 
over time. The narrative around IdrA still regards him as one of the greatest 
Zerg players of all time, at a time where not many players used Zerg in tournament
play. Thus, in trying to tease out balance issues in the game, IdrA serves a 
uniquely important role. 

(Maybe look into Nestea here too)


Note, Idra's story really captures the "long slump" well I think [](https://liquipedia.net/starcraft2/IdrA)


## Stephano's Story

```{r stephano_story}
elo_story('1996_Stephano') %>% qplot(x = id, y = elo, data = .) %>%
  add_time_mark(17425, "MLG Dallas 2011") %>%
  add_time_mark(28459, "ESWC France 2011") %>%
  add_time_mark(35275, "IEM VI Cologne 2011") %>%
  add_time_mark(41932, "ESWC France Finals 2011") %>%
  add_time_mark(47480, "Blizzard Cup 2011") %>%
  add_time_mark(49626, "eOSL Winter 2012") %>%
  add_time_mark(58402, "Lone Star Clash 2012") %>%
  add_time_mark(76359, "MLG Summer 2012")
```

Stephano doesn't have a reliable tournament presence prior to 2011. This is possibly because he started playing Starcraft under the handle Sat. Unfortunately, this set doesn't include games from that handle. 

Stephano's presence in the tournament scene seems to be marked by a slow, steady climb to top-tier play. By July 2011, Stephano had placed 1st in the Master Series, and by Oct 2011 he would go on and win the ESWC against MarineKing and MaNa. 

In November 2011, Stephano has a disappointing run in the Battle of Berlin, Dreamhack, and the Blizzard Cup. The model captures this run nicely, and Stephano quickly rebounds shortly after. 

By March 2012, Stephano has reached his prime. 


Stephano could be an interesting contrast to IdrA. But the key objection to contrasting him to IdrA is that you're comparing a largely French player to an American one. Those talent pools may be different, and thus their elos not comparable. In order to use Stephano as a contrast we need to demonstrate that his elo gives good expectations regarding his tournament wins/losses. 

Here's an idea - create a data set of before/after games just after a balance change. Among all players of a specific race, does average elo go up or down?


## MarineKing's Story

```{r mk_story}
elo_story('959_MarineKing') %>% qplot(x = id, y = elo, data = .) %>% 
  add_time_mark(3731, "GSL S2 2010") %>%
  add_time_mark(8598, "GSL Code S 2011") %>%
  add_time_mark(21707, "GSL Up 2011") %>%
  add_time_mark(33671, "GSL Code A 2011") %>%
  add_time_mark(41730, "GSL Up Oct 2011")
elo_story('959_MarineKing', df20) %>% qplot(x = id, y = elo, data = .)
elo_story('959_MarineKing', df32) %>% qplot(x = id, y = elo, data = .)
```

Made himself famous at GSL S2 with marine split stimming and 2-rax. 

Dec. 7, 2010, Blizzard puts him on a top 200 list. 


## Mvp's Story

```{r mvp_story}
elo_story('590_Mvp') %>% qplot(x = id, y = elo, data = .)
elo_story('590_Mvp', df20) %>% qplot(x = id, y = elo, data = .)
elo_story('590_Mvp', df32) %>% qplot(x = id, y = elo, data = .)
```

## Life's

```{r mvp_story}
elo_story('2747_Life') %>% qplot(x = id, y = elo, data = .)
elo_story('2747_Life', df20) %>% qplot(x = id, y = elo, data = .)
elo_story('2747_Life', df32) %>% qplot(x = id, y = elo, data = .)
```

## hG's Story

```{r mvp_story}
elo_story('1736_hG') %>% qplot(x = id, y = elo, data = .)
```



```{r}
df32 %>% 
  group_by(P1) %>% 
  summarise(high_score = max(elo_P1)) %>% arrange(desc(high_score)) %>% 
  head(20)
```